---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rcorex

<!-- badges: start -->
[![R build status](https://github.com/jpkrooney/rcorex/workflows/R-CMD-check/badge.svg)](https://github.com/jpkrooney/rcorex/actions)
<!-- badges: end -->

Total correlation explanation is an unsupervised learning algorithm for discovering structure in high dimensional data. Total correlation explanation has been implemented in Python as CorEx and related modules (https://github.com/gregversteeg/CorEx). The initial aim of rcorex is to implement Total Correlation Explanation in the R statistical software, specifically to replicate the functionality of the Bio CorEx Python module ( https://github.com/gregversteeg/bio_corex ).

The theoretical framework behind the CorEx and Bio CorEx Python modules are laid out in the following academic papers:  
1. [Discovering Structure in High-Dimensional Data Through Correlation Explanation](http://arxiv.org/abs/1406.1222)  
2. [Maximally Informative Hierarchical Representions of High-Dimensional Data](https://arxiv.org/abs/1410.7404)  
3. [Comprehensive discovery of subsample gene expression components by information explanation: therapeutic implications in cancer](https://bmcmedgenomics.biomedcentral.com/articles/10.1186/s12920-017-0245-6)  



## Installation
rcorex can be installed from Github:  

```{r gh-instal, eval = FALSE}
devtools::install_github("jpkrooney/rcorex")
```
