---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rcorex: Discover latent structure in high dimensional data. <img src='man/figures/rcorexlogo.svg' align="right" height="139" />


<!-- badges: start -->
[![R build status](https://github.com/jpkrooney/rcorex/workflows/R-CMD-check/badge.svg)](https://github.com/jpkrooney/rcorex/actions)
[![Codecov test coverage](https://codecov.io/gh/jpkrooney/rcorex/branch/master/graph/badge.svg)](https://codecov.io/gh/jpkrooney/rcorex?branch=master)
[![DOI](https://zenodo.org/badge/326518526.svg)](https://zenodo.org/badge/latestdoi/326518526)
<!-- badges: end -->


Total correlation explanation is method for discovering latent structure in high dimensional data. Total correlation explanation has been implemented in Python as CorEx and related modules (https://github.com/gregversteeg/CorEx). The initial aim of rcorex is to implement Total Correlation Explanation in the R statistical software, specifically to replicate the functionality of the BioCorEx Python module ( https://github.com/gregversteeg/bio_corex ) which is designed to  be suitable for biomedical datasets. This is implemented in the `biocorex()` command.

The theoretical framework behind the CorEx and Bio CorEx Python modules are laid out in the following academic papers:  
1. [Discovering Structure in High-Dimensional Data Through Correlation Explanation](http://arxiv.org/abs/1406.1222)  
2. [Maximally Informative Hierarchical Representions of High-Dimensional Data](https://arxiv.org/abs/1410.7404)  
3. [Comprehensive discovery of subsample gene expression components by information explanation: therapeutic implications in cancer](https://bmcmedgenomics.biomedcentral.com/articles/10.1186/s12920-017-0245-6)  


## Installation
rcorex can be installed from Github:  

```{r gh-instal, eval = FALSE}
# install.packages("remotes")
remotes::install_github("jpkrooney/rcorex")
```


## Minimal example
To fit a CorEx model in rcorex we can use the `biocorex()` command. `biocorex()` accepts a data.frame or a matrix as input, however as with the Python implementation of Bio CorEx, all variables must have the same data-type and currently only "discrete" or "gaussian" data are allowed as marginal descriptions, which apply to all columns.

```{r example}
library(rcorex)
# make a small dataset
df1 <- matrix(c(0,0,0,0,0,
             0,0,0,1,1,
             1,1,1,0,0,
             1,1,1,1,1), ncol=5, byrow = TRUE)

# fit biocorex
set.seed(1234)
fit1 <- biocorex(df1, n_hidden = 2, dim_hidden = 2, marginal_description = "discrete", logpx_method = "pycorex")
plot(fit1)
summary(fit1)

# What was the total correlation for each hidden dimension ?
fit1$tcs

# Which variables were clustered together?
fit1$clusters

# Which labels were assigned to each row of data for hidden cluster 1?
fit1$labels[, 1]

# And for hidden cluster 2?
fit1$labels[, 2]
```


## Hierarchical CorEx example
`rcorex` can search for hierarchical structure in data by using the labels output from an `rcorex` object as the input to the next layer in the hierarchy. This is shown in the following example using R's inbuilt `iris` dataset.

```{r iris_example_layer1}
library(rcorex)
library(ggraph)

set.seed(1234)

# Load iris dataset
data("iris")

# Need to convert species factor variable to indicator variables
iris <- data.frame(iris , model.matrix(~iris$Species)[,2:3])
iris$Species <- NULL

# fit first layer of CorEx 
layer1 <- biocorex(iris, 3, 2, marginal_description = "gaussian", repeats = 5, logpx_method = "pycorex")

```

Note the use of the `repeats = 5` argument to `biocorex`. This acts to run `biocorex` not once, but 5 times and `biocorex` automatically selects the run which produces the maximal TC to return to the user (unless the `return_all_runs` argument is set to `TRUE`).

We can then use the labels from `layer1` as the input for a second layer of CorEx to discover hierarchical structure. Note that the value used for n_hidden should be lower in the second layer than it was in the first.

```{r iris_example_layer2}

# fit second layer of CorEx
layer2 <- biocorex(layer1$labels, 1,2, marginal_description = "discrete", repeats = 5, logpx_method = "pycorex")

# make a network tidygraph of hierarchical layers
g_hier <- make_corex_tidygraph( list(layer1, layer2))

# Plot network graph of hierarchical layers
ggraph(g_hier, layout = "fr") +
    geom_node_point(aes(size = node_size), show.legend = FALSE) +
    geom_edge_hive(aes(width = thickness), alpha = 0.75, show.legend = FALSE) +
    scale_edge_width(range = c(0.3, 3)) +
    geom_node_text(aes(label = names), repel = TRUE) +
    theme_graph()

```

Additional hierarchical layers can be identified in larger datasets. 


## Citation
If you find `rcorex` useful in academic work please cite the package as follows:
```{r citation}
citation("rcorex")
```


